Traffic Aware Resource Allocation Schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} Systems

### 2015-03-19 Report/joint_wsrm_formulation.tex
with_markku_corrections

The queue minimizing algorithms are discussed extensively in the networking literature to provide congestion-free routing between any two nodes in the network. One such algorithm  is the \emph{backpressure algorithm} \cite{tassiulas,georgiadis2006resource,neely2010stochastic}. It determines an optimal control policy in the form of rate or resource allocation for the nodes in the network by considering the differential backlogged packets between the source and the destination nodes. Even though the algorithm is primarily designed for the wired infrastructure, it can be extended to the wireless networks by designing the user rate variable \me{t_k} in accordance to the wireless network.

@@

To minimize the total number of backlogged packets for an instant \me{i}, the optimal transmission rate of all users are obtained by minimizing the Lyapunov function drift expressed as  
\iftoggle{single_column}{
\begin{equation}\label{eqn-3.1}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} = \frac{1}{2} \Big [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \Big ].
\end{equation}}{
\begin{multline}\label{eqn-3.1}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} = \\ \frac{1}{2} \Big [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \Big ].
\end{multline}}
In order to eliminate the nonlinear operator \me{[x]^+}, we bound the expression in \eqref{eqn-3.1} as
\begin{equation}
\leq \sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} + \sum_{k \in \mc{U}} Q_k(i) \set{ \lambda_k(i) - t_k(i) }
\label{drift-exp}
\end{equation}
by using the following inequality 
\begin{equation}
\sset{\max(Q-t,0) + \lambda)}^2 \leq Q^2 + t^2 + \lambda^2 + 2 Q (\lambda - t).
\end{equation}
The total number of backlogged packets at any given instant \me{i} is reduced by minimizing the conditional expectation of the Lyapunov drift expression \eqref{drift-exp} given the current number of queued packets \me{\mbfa{Q}(i)} waiting in the system. The expectation is taken over all possible arrival and transmission rates of the users to obtain the optimal rate allocation strategy.

@@

The expression in \eqref{backpressure-approx} looks similar to the \ac{WSRM} formulation if the weights in the \ac{WSRM} problem are replaced by the number of backlogged packets corresponding to the users. The above discussed approach is extended for the wireless networks in \cite{weeraddana2011resource}, in which the queues are used as weights in the \ac{WSRM} formulation to determine the transmit precoders. Since the expectation is minimized by minimizing the function inside, the \ac{Q-WSRM} formulation is given by 
\begin{IEEEeqnarray}{LCl} \label{q_gen_sum}
\underset{\substack{\mvec{m}{l,k,n}, \mvec{w}{l,k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, Q_k \left ( \sum_{n = 1}^N \sum_{l=1}^L \log_2(1 + \gamma_{l,k,n}) \right ) \IEEEyessubnumber \label{eqn-3.1.1} \\
\text{subject to.} & \quad & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}} \fall b. \IEEEyessubnumber \eqspace \label{eqn-3.1.3}
\end{IEEEeqnarray}

To avoid excessive allocation of the resources, we include an additional rate constraint \me{t_k \leq Q_k} to address \me{[x]^+} operation in \eqref{eqn-2a}. The rate constrained version of the \ac{Q-WSRM}, denoted by \ac{Q-WSRME} problem for a cellular system, is given by \label{q_gen_sum-1} with the additional constraints as
\begin{IEEEeqnarray}{rCl} \label{eqn-3.1.4}
\sum_{n=1}^N \sum_{l = 1}^L \log_2(1 + \gamma_{l,k,n}) \leq Q_k \: \fall k \in \mc{U}
\end{IEEEeqnarray}
where the precoders are associated with \me{\gamma_{l,k,n}} defined in \eqref{eq:SINR}. By using the number of queued packets as the weights, the resources can be allocated to the user with the more backlogged packets, which essentially does greedy allocation.

>>>>

The queue minimizing algorithms have been extensively discussed in the networking literature to provide congestion-free routing between any two nodes in the network. One such algorithm  is the \emph{backpressure algorithm} \cite{tassiulas,georgiadis2006resource,neely2010stochastic}. It determines an optimal control policy in the form of rate or resource allocation for the nodes in the network by considering the differential backlogged packets between the source and the destination nodes. Even though the algorithm is primarily designed for the wired infrastructure, it can be extended to the wireless networks by designing the user rate variable \me{t_k} in accordance to the wireless network.

@@

To minimize the total number of backlogged packets for time instant \me{i}, the optimal transmission rate of all users are obtained by minimizing the Lyapunov drift expressed as  
\iftoggle{single_column}{
\begin{equation}\label{eqn-3.1}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} = \frac{1}{2} \Big [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \Big ].
\end{equation}}{
\begin{multline}\label{eqn-3.1}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} = \\ \frac{1}{2} \Big [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \Big ].
\end{multline}}
In order to eliminate the nonlinear operator \me{[x]^+}, we bound the expression in \eqref{eqn-3.1} as
\begin{equation}
\leq \sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} + \sum_{k \in \mc{U}} Q_k(i) \set{ \lambda_k(i) - t_k(i) }
\label{drift-exp}
\end{equation}
by using the following inequality 
\begin{equation}
\sset{\max(Q-t,0) + \lambda)}^2 \leq Q^2 + t^2 + \lambda^2 + 2 Q (\lambda - t).
\end{equation}
The total number of backlogged packets at any given instant \me{i} is reduced by minimizing the conditional expectation of the Lyapunov drift expression \eqref{drift-exp} given the current number of queued packets \me{\mbfa{Q}(i)} waiting in the system. The expectation is taken over all possible arrival and transmission rates of the users to obtain the optimal rate allocation strategy.

@@

The expression in \eqref{backpressure-approx} looks similar to the \ac{WSRM} formulation if the weights in the \ac{WSRM} problem are replaced by the numbers of backlogged packets of the corresponding users. The above approach was extended for the wireless networks in \cite{weeraddana2011resource}, in which the queues were used as weights in the \ac{WSRM} formulation to determine the transmit precoders. Since the expectation is minimized by minimizing the function inside, the \ac{Q-WSRM} formulation is given by 
\begin{IEEEeqnarray}{LCl} \label{q_gen_sum}
\underset{\substack{\mvec{m}{l,k,n}, \mvec{w}{l,k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, Q_k \left ( \sum_{n = 1}^N \sum_{l=1}^L \log_2(1 + \gamma_{l,k,n}) \right ) \IEEEyessubnumber \label{eqn-3.1.1} \\
\text{subject to.} & \quad & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}} \fall b. \IEEEyessubnumber \eqspace \label{eqn-3.1.3}
\end{IEEEeqnarray}

To avoid excessive allocation of the resources, we include an additional rate constraint \me{t_k \leq Q_k} to address \me{[x]^+} operation in \eqref{eqn-2a}. The rate constrained version of the \ac{Q-WSRM}, denoted by \ac{Q-WSRME} problem for a cellular system, is given by \label{q_gen_sum-1} with the additional constraints as
\begin{IEEEeqnarray}{rCl} \label{eqn-3.1.4}
\sum_{n=1}^N \sum_{l = 1}^L \log_2(1 + \gamma_{l,k,n}) \leq Q_k \: \fall k \in \mc{U}
\end{IEEEeqnarray}
where the precoders are associated with \me{\gamma_{l,k,n}} defined in \eqref{eq:SINR}. By using the number of queued packets as the weights, the resources can be allocated to the user with more backlogged packets; this essentially results in greedy allocation.

### 2015-03-19 Report/kkt_mse_approach.tex
with_markku_corrections

In this section, we discuss an alternative way to decentralize the precoder design across the coordinating \acp{BS} in \me{\mc{B}} based on the \ac{MSE} reformulation method discussed in Section \ref{sec-3.3}. In contrast to Section \ref{sec-4.1} and \ref{sec-4.2}, the problem is solved using the \ac{KKT} conditions in which the transmit precoders, receive beamformers and the subgradient updates are performed at the same instant to minimize the global queue deviation objective with few number of iterations. The proposed methods in this section provide algorithms that can be of practical importance owing to the limited signaling requirements. \review{We consider an idealized \ac{TDD} system due to the knowledge of complete channel information at the transmitter}.  Similar work has been considered for the \ac{WSRM} problem with minimum rate constraints in \cite{kaleva2013decentralized,kaleva2013primal}. Since the formulation in \cite{kaleva2013decentralized,kaleva2013primal} are similar to the \ac{Q-WSRME} scheme with an additional maximum rate constraint \eqref{eqn-3.1.4}, it requires explicit dual variables to handle the maximum rate constraint, thereby making the problem difficult to solve in an iterative manner.

@@

With the assumption of either one of the above conditions to be true, the problem in \eqref{eqn-mse-2} can be written as
\iftoggle{single_column}{\allowdisplaybreaks
\begin{IEEEeqnarray}{rCl}\label{kkt-mse-1}
\underset{\substack{t_{l,k,n},\mvec{m}{l,k,n}, \\ \epsilon_{l,k,n}, \mvec{w}{l,k,n}}}{\text{minimize}} &\hspace{0.5cm}& \sum_{b \in \mc{B}} \sum_{k \in \mc{U}_b} \, a_k \, \Big ( Q_k - \sum_{n = 1}^N \sum_{l = 1}^{L} t_{l,k,n} \Big )^q \IEEEyessubnumber \label{kkt-mse-1.1} \\
\text{subject to} & \hspace{0.5cm} & \nonumber \\
\alpha_{l,k,n} : & \hspace{0.5cm} & \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 +  \enoise + \sum_{\mathclap{(x,y) \neq (l,k)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_y,k,n} \mvec{m}{x,y,n} \right |^2 \leq \epsilon_{l,k,n} \eqspace \IEEEyessubnumber \label{kkt-mse-1.2} \\
\sigma_{l,k,n} : & \hspace{0.5cm} & \log(\tilde{\epsilon}_{l,k,n}) + \frac{\left ( {\epsilon}_{l,k,n} - \tilde{\epsilon}_{l,k,n} \right ) }{\tilde{\epsilon}_{l,k,n}} \leq -t_{l,k,n} \, \log(2) \IEEEyessubnumber \label{kkt-mse-1.3} \\
\delta_b : & \hspace{0.5cm} & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l = 1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}}\; \fall b, \IEEEyessubnumber \label{kkt-mse-1.4}
\end{IEEEeqnarray}
}{
\begin{IEEEeqnarray}{Rl}\label{kkt-mse-1}
	\underset{\substack{t_{l,k,n},\mvec{m}{l,k,n}, \\ \epsilon_{l,k,n}, \mvec{w}{l,k,n}}}{\text{minimize}} & \quad \sum_{b \in \mc{B}} \sum_{k \in \mc{U}_b} \, a_k \, \Big ( Q_k - \sum_{n = 1}^N \sum_{l = 1}^{L} t_{l,k,n} \Big )^q \IEEEyessubnumber \label{kkt-mse-1.1} \\
	\text{subject to} & \nonumber \\
	\alpha_{l,k,n} : &  \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 +  \enoise \nonumber \\
	& + \sum_{\mathclap{(x,y) \neq (l,k)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_y,k,n} \mvec{m}{x,y,n} \right |^2 \leq \epsilon_{l,k,n} \IEEEyessubnumber \label{kkt-mse-1.2} \\
	\sigma_{l,k,n} : &  \log_2(\tilde{\epsilon}_{l,k,n}) + \frac{\left ( {\epsilon}_{l,k,n} - \tilde{\epsilon}_{l,k,n} \right ) }{\log(2)\tilde{\epsilon}_{l,k,n}} \leq -t_{l,k,n} \IEEEyessubnumber \label{kkt-mse-1.3} \\
	\delta_b : & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l = 1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}} \; \fall b \IEEEyessubnumber \eqspace \label{kkt-mse-1.4}
\end{IEEEeqnarray}
}
where \me{\alpha_{l,k,n},\sigma_{l,k,n}} and \me{\delta_b} are the dual variables corresponding to the constraints defined in \eqref{kkt-mse-1.2}, \eqref{kkt-mse-1.3} and \eqref{kkt-mse-1.4}.

@@

}{{\allowdisplaybreaks

@@

\end{IEEEeqnarray}}}
Since the dual variables \me{\alpha^{(i)}} and \me{\sigma^{(i)}} are interdependent in \eqref{kkt-mse-4}, one has to be fixed to optimize for the other. So, \me{\alpha^{(i)}} is fixed to evaluate \me{\sigma^{(i)}} using \eqref{kkt-mse-4}. \review{At each iteration, the dual variables \me{\alpha^{(i)}} are linearly interpolated with any point between the fixed iterate \me{\alpha^{(i-1)}} and \me{\tfrac{\sigma^{(i)}}{\epsilon^{(i)}}} using a step size \me{\rho \in (0,1)}. The choice of \me{\rho} depends on the system model and it affects the convergence behavior. It reduces the oscillations in the objective function when \me{\sigma^{(i)}} is negative due to over allocation.}

When the allocated rate \me{t_k^{(i-1)}} is greater than the number of queued packets \me{Q_k} for a user \me{k}, the corresponding dual variable \me{\sigma^{(i)}} will be negative and due to the projection operator \me{[x]^+} in \eqref{kkt-mse-4.2}, it will be zero, thereby forcing \me{\alpha_k^{(i)} < \alpha_k^{(i-1)}} as in \eqref{kkt-mse-4.1}. Once the \me{\alpha_k^{(i)}} is reduced, the precoder weight in \eqref{kkt-mse-4.3} is lowered to make the rate \me{t_k^{(i)} < t_k^{(i-1)}} eventually. \review{The choice of \me{\rho} is susceptible to the system model under consideration, which affects the convergence speed of the iterative algorithm. In all simulations, the step size \me{\rho} is fixed to \me{0.1} irrespectively.}

@@

To avoid the back-haul exchanges between \acp{BS}, as an alternative approach, users may perform all processing required and \acp{BS} will update the precoders based on the feedback information from the users. Upon receiving the transmit precoders from \acp{BS}, each user will update the receive beamformer \me{\mvec{w}{l,k,n}}, the \ac{MSE} \me{\epsilon_{l,k,n}}, and the dual variables \me{\lambda_{l,k,n}} and \me{\alpha_{l,k,n}}. The updated \me{\alpha_{l,k,n}} and \me{\mvec{w}{l,k,n}} are notified to the \acp{BS} using two separate precoded uplink pilot symbols with \me{\tilde{\mbf{w}}_{l,k,n}^{(i)} = \sqrt{\alpha_{l,k,n}^{(i)}}\mvec{w}{l,k,n}^{\ast(i)}} and \me{\bar{\mbf{w}}_{l,k,n}^{(i)} = \alpha_{l,k,n}^{(i)}\mvec{w}{l,k,n}^{\ast(i)}} as the precoders. On receiving the precoded uplink pilots, each \ac{BS} use the effective channel \me{\mvec{H}{b,k,n}^{\mathrm{T}} \tilde{\mbf{w}}_{l,k,n}^{(i)}} and \me{\mvec{H}{b,k,n}^{\mathrm{T}} \bar{\mbf{w}}_{l,k,n}^{(i)}} in \eqref{kkt-mse-4.3} to update the transmit precoders, where \me{\mbf{x}^{\ast}} is the complex conjugate of \me{\mbf{x}}. Finally, Algorithm \ref{algo-4} outlines the distributed precoder design using the \ac{KKT} based \ac{MSE} reformulated \ac{JSFRA} problem.
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=1}, \me{\mbf{w}^{(0)}_{l,k,n}, \tilde{\epsilon}_{l,k,n}} randomly, dual variables \me{{\alpha}_{l,k,n}^{(0)} = 1}, and \me{I_{\max}} for certain value}
 \ForEach{\ac{BS} \me{b \in \mc{B}}}{
 initialize \me{i = 0} \;
 \Repeat{until convergence or \me{i \geq I_{\max}}}{
	update \me{\mvec{m}{l,k,n}^{(i)}} using \eqref{kkt-mse-4.3}, and perform downlink transmission \;
	find \me{\mvec{w}{l,k,n}^{(i)}} using \eqref{kkt-mse-4.6} at each user \;
	evaluate \me{\epsilon_{l,k,n}^{(i)}}, \me{t_{l,k,n}^{(i)}}, \me{\sigma_{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} using \eqref{kkt-mse-4.4} and \eqref{kkt-mse-4.5}, \eqref{kkt-mse-4.2} and \eqref{kkt-mse-4.1} at each user with the updated \me{\mvec{w}{l,k,n}^{(i)}} \;
	using precoded uplink pilots, \me{\mvec{m}{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} are notified to all \acp{BS} in \me{\mathcal{B}} \;
	\me{i = i + 1} \;
 }
 }
 \caption{\ac{KKT} approach for the \ac{JSFRA} scheme}
 \label{algo-4}
\end{algorithm}

\review{Algorithm \ref{algo-4} outlines a practical way of updating the transmit and the receive beamformers by using \ac{OTA} signaling with the precoded pilots. Unlike \ac{PD} or \ac{ADMM} schemes, all variables are updated at once, \textit{i.e}, the \ac{SCA} update of \me{\epsilon^{(i-1)}}, \ac{AO} update of \me{\mvec{w}{l,k,n}} and the dual variable update of \me{\alpha} using subgradient. It is not guaranteed to obtain the same point as that of the centralized problem. The algorithm in \eqref{kkt-mse-4} is identical to \eqref{eqn-mse-2}, if the receivers \me{\mvec{w}{l,k,n}} and the \ac{MSE} operating point \me{\epsilon_{l,k,n}^{(i-1)}} are fixed to find the optimal transmit precoders \me{\mvec{m}{l,k,n}} and the dual variable \me{\alpha_{l,k,n}}. Note that it requires four nested loops to obtain the centralized solution, namely, the receive beamformer loop, \ac{MSE} operating point loop, dual variable update loop and the bisection method to find the transmit precoders. To avoid this, the proposed method performs group update of all variables at once to obtain the transmit and the receive beamformers with the limited number of iterations, thus it achieves improved speed of convergence. Even though the convergence is not theoretically guaranteed, it converges in all numerical experiments considered in Section \ref{sec-5.2}. If the step size \me{\rho < 1}, the algorithm for \me{\ell_2} norm will converge by using the arguments on controlling overallocation and the nonincreasing objective function, since the earlier iterates are the operating point for the current iteration.}

>>>>

In this section, we discuss an alternative way to decentralize the precoder design across the coordinating \acp{BS} in \me{\mc{B}} based on the \ac{MSE} reformulation method discussed in Section \ref{sec-3.3}. In contrast to Section \ref{sec-4.1} and \ref{sec-4.2}, the problem is solved using the \ac{KKT} conditions in which the transmit precoders, receive beamformers and the subgradient updates are performed at the same instant to minimize the global queue deviation objective with few number of iterations. The proposed methods in this section provide algorithms that can be of practical importance owing to the limited signaling requirements. \review{We consider an idealized \ac{TDD} system due to the knowledge of complete channel information at the transmitter}. Similar work has been considered for the \ac{WSRM} problem with minimum rate constraints in \cite{kaleva2013decentralized,kaleva2013primal}. Since the formulations in \cite{kaleva2013decentralized,kaleva2013primal} are similar to that of the \ac{Q-WSRME} scheme with an additional maximum rate constraint \eqref{eqn-3.1.4}, it requires explicit dual variables to handle the maximum rate constraint, thereby making the problem difficult to solve in an iterative manner.

@@

With the assumption of either one of the above conditions to be true, the problem in \eqref{eqn-mse-2} can be written as
\allowdisplaybreaks\iftoggle{single_column}{
\begin{IEEEeqnarray}{rCl}\label{kkt-mse-1}
\underset{\substack{t_{l,k,n},\mvec{m}{l,k,n}, \\ \epsilon_{l,k,n}, \mvec{w}{l,k,n}}}{\text{minimize}} &\hspace{0.5cm}& \sum_{b \in \mc{B}} \sum_{k \in \mc{U}_b} \, a_k \, \Big ( Q_k - \sum_{n = 1}^N \sum_{l = 1}^{L} t_{l,k,n} \Big )^q \IEEEyessubnumber \label{kkt-mse-1.1} \\
\text{subject to} & \hspace{0.5cm} & \nonumber \\
\alpha_{l,k,n} : & \hspace{0.5cm} & \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 +  \enoise + \sum_{\mathclap{(x,y) \neq (l,k)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_y,k,n} \mvec{m}{x,y,n} \right |^2 \leq \epsilon_{l,k,n} \eqspace \IEEEyessubnumber \label{kkt-mse-1.2} \\
\sigma_{l,k,n} : & \hspace{0.5cm} & \log(\tilde{\epsilon}_{l,k,n}) + \frac{\left ( {\epsilon}_{l,k,n} - \tilde{\epsilon}_{l,k,n} \right ) }{\tilde{\epsilon}_{l,k,n}} \leq -t_{l,k,n} \, \log(2) \IEEEyessubnumber \label{kkt-mse-1.3} \\
\delta_b : & \hspace{0.5cm} & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l = 1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}}\; \fall b, \IEEEyessubnumber \label{kkt-mse-1.4}
\end{IEEEeqnarray}
}{\begin{IEEEeqnarray}{Rl}\label{kkt-mse-1}
	\underset{\substack{t_{l,k,n},\mvec{m}{l,k,n}, \\ \epsilon_{l,k,n}, \mvec{w}{l,k,n}}}{\text{minimize}} & \quad \sum_{b \in \mc{B}} \sum_{k \in \mc{U}_b} \, a_k \, \Big ( Q_k - \sum_{n = 1}^N \sum_{l = 1}^{L} t_{l,k,n} \Big )^q \IEEEyessubnumber \label{kkt-mse-1.1} \\
	\text{subject to} & \nonumber \\
	\alpha_{l,k,n} : &  \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 +  \enoise \nonumber \\
	& + \sum_{\mathclap{(x,y) \neq (l,k)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_y,k,n} \mvec{m}{x,y,n} \right |^2 \leq \epsilon_{l,k,n} \IEEEyessubnumber \label{kkt-mse-1.2} \\
	\sigma_{l,k,n} : &  \log_2(\tilde{\epsilon}_{l,k,n}) + \frac{\left ( {\epsilon}_{l,k,n} - \tilde{\epsilon}_{l,k,n} \right ) }{\log(2)\tilde{\epsilon}_{l,k,n}} \leq -t_{l,k,n} \IEEEyessubnumber \label{kkt-mse-1.3} \\
	\delta_b : & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l = 1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}} \; \fall b \IEEEyessubnumber \eqspace \label{kkt-mse-1.4}
\end{IEEEeqnarray}}
where \me{\alpha_{l,k,n},\sigma_{l,k,n}} and \me{\delta_b} are the dual variables corresponding to the constraints defined in \eqref{kkt-mse-1.2}, \eqref{kkt-mse-1.3} and \eqref{kkt-mse-1.4}.

@@

}{\allowdisplaybreaks

@@

\end{IEEEeqnarray}}
Since the dual variables \me{\alpha^{(i)}} and \me{\sigma^{(i)}} are interdependent in \eqref{kkt-mse-4}, one has to be fixed to optimize for the other. So, \me{\alpha^{(i)}} is fixed to evaluate \me{\sigma^{(i)}} using \eqref{kkt-mse-4}. \review{At each iteration, the dual variables \me{\alpha^{(i)}} are linearly interpolated with any point between the fixed iterate \me{\alpha^{(i-1)}} and \me{\tfrac{\sigma^{(i)}}{\epsilon^{(i)}}} using a step size \me{\rho \in (0,1)}. The choice of \me{\rho} depends on the system model and it affects the convergence behavior. It reduces the oscillations in the objective function when \me{\sigma^{(i)}} is negative due to over-allocation.}

When the allocated rate \me{t_k^{(i-1)}} is greater than the number of queued packets \me{Q_k} for a user \me{k}, the corresponding dual variable \me{\sigma^{(i)}} will be negative and due to the projection operator \me{[x]^+} in \eqref{kkt-mse-4.2}, it will be zero, thereby forcing \me{\alpha_k^{(i)} < \alpha_k^{(i-1)}} as in \eqref{kkt-mse-4.1}. Once \me{\alpha_k^{(i)}} is reduced, the precoder weight in \eqref{kkt-mse-4.3} is lowered to make the rate \me{t_k^{(i)} < t_k^{(i-1)}} eventually. \review{The choice of \me{\rho} is susceptible to the system model under consideration, which affects the convergence speed of the iterative algorithm. In all simulations, the step size \me{\rho} is fixed to \me{0.1} irrespectively.}

@@

To avoid the back-haul exchanges between the \acp{BS}, as an alternative approach, users perform all processing required and \acp{BS} will update the precoders based on the feedback information from the users. Upon receiving the transmit precoders from \acp{BS}, each user will update the receive beamformer \me{\mvec{w}{l,k,n}}, the \ac{MSE} \me{\epsilon_{l,k,n}}, and the dual variables \me{\lambda_{l,k,n}} and \me{\alpha_{l,k,n}}. The updated \me{\alpha_{l,k,n}} and \me{\mvec{w}{l,k,n}} are notified to the \acp{BS} using two separate precoded uplink pilot symbols with \me{\tilde{\mbf{w}}_{l,k,n}^{(i)} = \sqrt{\alpha_{l,k,n}^{(i)}}\mvec{w}{l,k,n}^{\ast(i)}} and \me{\bar{\mbf{w}}_{l,k,n}^{(i)} = \alpha_{l,k,n}^{(i)}\mvec{w}{l,k,n}^{\ast(i)}} as the precoders. On receiving the precoded uplink pilots, each \ac{BS} use the effective channel \me{\mvec{H}{b,k,n}^{\mathrm{T}} \tilde{\mbf{w}}_{l,k,n}^{(i)}} and \me{\mvec{H}{b,k,n}^{\mathrm{T}} \bar{\mbf{w}}_{l,k,n}^{(i)}} in \eqref{kkt-mse-4.3} to update the transmit precoders, where \me{\mbf{x}^{\ast}} is the complex conjugate of \me{\mbf{x}}. Finally, Algorithm \ref{algo-4} outlines the distributed precoder design using the \ac{KKT} based \ac{MSE} reformulated \ac{JSFRA} problem.
\begin{algorithm}
 \SetAlgoLined
 \DontPrintSemicolon
 \BlankLine
 \SetKwInput{KwInit}{Initialize}
 \KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
 \KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
 \KwInit{\me{i=1}, \me{\mbf{w}^{(0)}_{l,k,n}, \tilde{\epsilon}_{l,k,n}} randomly, dual variables \me{{\alpha}_{l,k,n}^{(0)} = 1}, and \me{I_{\max}} for certain value}
 \ForEach{\ac{BS} \me{b \in \mc{B}}}{
 initialize \me{i = 0} \;
 \Repeat{until convergence or \me{i \geq I_{\max}}}{
	update \me{\mvec{m}{l,k,n}^{(i)}} using \eqref{kkt-mse-4.3}, and perform downlink transmission \;
	find \me{\mvec{w}{l,k,n}^{(i)}} using \eqref{kkt-mse-4.6} at each user \;
	evaluate \me{\epsilon_{l,k,n}^{(i)}}, \me{t_{l,k,n}^{(i)}}, \me{\sigma_{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} using \eqref{kkt-mse-4.4} and \eqref{kkt-mse-4.5}, \eqref{kkt-mse-4.2} and \eqref{kkt-mse-4.1} at each user with the updated \me{\mvec{w}{l,k,n}^{(i)}} \;
	using precoded uplink pilots, \me{\mvec{m}{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} are notified to all \acp{BS} in \me{\mathcal{B}} \;
	\me{i = i + 1} \;
 }
 }
 \caption{\ac{KKT} approach for the \ac{JSFRA} scheme}
 \label{algo-4}
\end{algorithm}

\review{Algorithm \ref{algo-4} outlines a practical way of updating the transmit and the receive beamformers by using \ac{OTA} signaling with the precoded pilots. Unlike \ac{PD} or \ac{ADMM} schemes, all variables are updated at once, \textit{i.e}, the \ac{SCA} update of \me{\epsilon^{(i-1)}}, the \ac{AO} update of \me{\mvec{w}{l,k,n}} and the dual variable update of \me{\alpha} using subgradient. It is not guaranteed to obtain the same point as that of the centralized problem. The algorithm in \eqref{kkt-mse-4} is identical to \eqref{eqn-mse-2}, if the receivers \me{\mvec{w}{l,k,n}} and the \ac{MSE} operating point \me{\epsilon_{l,k,n}^{(i-1)}} are fixed to find the optimal transmit precoders \me{\mvec{m}{l,k,n}} and the dual variable \me{\alpha_{l,k,n}}. Note that it requires four nested loops to obtain the centralized solution, namely, the receive beamformer loop, \ac{MSE} operating point loop, the dual variable update loop and the bisection method to find the transmit precoders. To avoid this, the proposed method performs group update of all variables at once to obtain the transmit and the receive beamformers with the limited number of iterations, thus it achieves improved speed of convergence. Even though the convergence is not theoretically guaranteed, it converges in all numerical experiments considered in Section \ref{sec-5.2}. If the step size \me{\rho < 1}, the algorithm for \me{\ell_2} norm will converge by using the arguments on controlling overallocation and the nonincreasing objective function, since the earlier iterates are the operating point for the current iteration.}

### 2015-03-19 Report/primal_approach.tex
with_markku_corrections

In \acl{PD}, the convex problem in \eqref{eqn-decent-1} is solved for the optimal transmit precoders in an iterative manner for a fixed \ac{BS} specific interference terms \me{\mbfa{\zeta}_{b_k}} using master-slave model \cite{pennanen2011decentralized}. The slave subproblem is solved in each \ac{BS} for the optimal transmit precoders only for the associated users by assuming fixed interference terms \me{\mbfa{\zeta}_{b_k}^{(i)}} in each \me{\ith{i}} iteration. Upon finding the optimal associated transmit precoders by each slave subproblems, the master problem is used to update the \ac{BS} specific interference terms \me{\mbfa{\zeta}_{b_k}^{(i+1)}} for the next iteration by using dual variables corresponding to the interference constraint \eqref{eqn-decent-3} as discussed in \cite{pennanen2011decentralized}. In this manner, the interference variables are updated until the global consensus is obtained. The primal approach is similar to the minimum power precoder design presented in \cite{pennanen2011decentralized}. Note that the master problem treats \me{\mbfa{\zeta}_{b}} as a variable and the slave subproblems assumes it to be a constant for each iteration to find the transmit precoders.

>>>>

In the \acl{PD}, the convex problem in \eqref{eqn-decent-1} is solved for the optimal transmit precoders in an iterative manner for a fixed \ac{BS} specific interference term \me{\mbfa{\zeta}_{b_k}} using the master-slave model \cite{pennanen2011decentralized}. The slave subproblem is solved in each \ac{BS} for the optimal transmit precoders only for the associated users by assuming fixed interference terms \me{\mbfa{\zeta}_{b_k}^{(i)}} in each \me{\ith{i}} iteration. Upon finding the optimal associated transmit precoders by each slave subproblems, the master problem is used to update the \ac{BS} specific interference terms \me{\mbfa{\zeta}_{b_k}^{(i+1)}} for the next iteration by using dual variables corresponding to the interference constraint \eqref{eqn-decent-3} as discussed in \cite{pennanen2011decentralized}. In this manner, the interference variables are updated until the global consensus is obtained. The primal approach is similar to the minimum power precoder design presented in \cite{pennanen2011decentralized}. Note that the master problem treats \me{\mbfa{\zeta}_{b}} as a variable and the slave subproblems assumes it to be a constant for each iteration to find the transmit precoders.
